Article 1:	Construct Validity and Correspondence of Google Trends

### 2022-02-19 paper1-cleaning.Rmd
write initial data and methods

vacc_hes_gtrends <- find_cross_sectional_trends(c('covid conspiracy',
                                                  'COVID-19 vaccine',
                                                  'Coronavirus',
                                                  'Covid-19'
                                                  # URLdecode('%2Fg%2F11j8_9sv06'),
                                                  # URLdecode('%2Fg%2F11j2cc_qll')
                                                  ), 
                            '2021-03-03', 12) %>%   
  pivot_wider(names_from = keyword, values_from = hits)  %>% 
  select(-date) %>% 
  janitor::clean_names()

@@

mask_gtrends <- find_cross_sectional_trends(c('Face mask',
                                              'Mask',
                                              # 'Civil and political rights',
                                              'Cloth face mask'
                                              ),
                                            '2020-07-02', 12)%>%   
  pivot_wider(names_from = keyword, values_from = hits)  %>% 
  select(-date) %>% 
  janitor::clean_names()

@@

@@

@@

```{r suicide}
suicide_measures_both <- read_delim(here('data','suicide',
                                    'Multiple Cause of Death, 1999-2020.txt')) %>% 
  janitor::clean_names() %>% 
  filter(is.na(notes)) %>% 
  select(fips = county_code, county, year = year_code, deaths_both = deaths, population)

suicide_measures_mcd <- read_delim(here('data','suicide',
                                    'Multiple Cause of Death, 1999-2020 only multiple cause.txt')) %>% 
  janitor::clean_names() %>% 
  filter(is.na(notes)) %>% 
  select(fips = county_code, county, year = year_code, deaths_mcd = deaths)

breaks <-  c(10,20,40,60,100,200,400,700,1300)
suicide_measures <- suicide_measures_both %>% 
  left_join(suicide_measures_mcd, by = c('fips', 'county', 'year')) %>% 
  mutate(deaths_mcd = na_if(deaths_mcd,'Missing'),
         deaths_both = na_if(deaths_both,'Missing'),
         deaths_mcd_supp = ifelse(deaths_mcd == 'Suppressed', '(0:9]', NA), 
         deaths_both_supp = ifelse(deaths_both == 'Suppressed', '(0:9]', NA), 
         deaths_mcd = na_if(deaths_mcd,'Suppressed'),
         deaths_both = na_if(deaths_both,'Suppressed'),
         deaths_mcd_cut = as.character(cut(as.numeric(deaths_mcd), breaks = breaks)),
         deaths_both_cut = as.character(cut(as.numeric(deaths_both), breaks = breaks)),
         deaths_mcd_cut = ifelse(!is.na(deaths_mcd_supp),
                                 deaths_mcd_supp,
                                 deaths_mcd_cut), 
         deaths_both_cut = ifelse(!is.na(deaths_both_supp),
                                  deaths_both_supp,
                                  deaths_both_cut)) %>% 
        select(fips, county, year, deaths_both_cut) %>% 
  mutate(deaths = factor(deaths_both_cut, ordered = T),
         deaths = fct_relevel(deaths, c("(0:9]",
                                        "(10,20]",
                                        "(20,40]", 
                                        "(40,60]",
                                        "(60,100]",    
                                        "(100,200]",
                                        "(200,400]",
                                        "(400,700]",   
                                        "(700,1.3e+03]")))
suicide_measures %>% 
  count(year, deaths) %>%
  ggplot(aes(x = deaths, y = n)) +
  geom_bar(stat = 'identity') +
  facet_wrap(~year)

@@

@@

suicide %>% 
  filter(fips %in% c("08003", '55057', '51059', '39085', 
           # '36119',
           '34021', '12071', '04017')) %>% 
  mutate(`actual deaths` = as.numeric(deaths) * 10) %>% 
  pivot_longer(cols = c('suicide', 'depression', 'suicide_hotline', 'actual deaths')) %>% 
  ggplot(aes(x = year, y = value, group = name, color = name)) +
  geom_line() + 
  geom_point() +
  facet_wrap(~ glue('{county}'), scale = 'free_y') + 
  theme_minimal()

>>>>

vacc_hes_gtrends <- find_cross_sectional_trends(c('covid conspiracy',
                                                  'COVID-19 vaccine',
                                                  'Coronavirus',
                                                  'Covid-19'
                                                  # URLdecode('%2Fg%2F11j8_9sv06'),
                                                  # URLdecode('%2Fg%2F11j2cc_qll')
                                                  ), 
                            '2021-05-26', 12) %>%   
  pivot_wider(names_from = keyword, values_from = hits)  %>% 
  select(-date) %>% 
  janitor::clean_names()

@@

mask_gtrends <- find_cross_sectional_trends(c('Face mask',
                                              'Mask',
                                              'Cloth face mask'
                                              ),
                                            '2020-07-02', 12)%>%   
  pivot_wider(names_from = keyword, values_from = hits)  %>% 
  select(-date) %>% 
  janitor::clean_names()

@@

corona_gtrend_2020 <- find_year_trends('Coronavirus', 2020) 
corona_gtrend_2021 <- find_year_trends('Coronavirus', 2021) 
corona_gtrend  <- corona_gtrend_2020 %>% bind_rows(corona_gtrend_2021)

@@

covid <- covid %>% 
  left_join(corona_gtrend, by = c('DMA_google', 'date')) %>% 
  group_by(fips) %>%
  mutate(covid_rate  = na.approx(covid_rate, na.rm = FALSE),
         covid_19    = na.approx(covid_19, na.rm = FALSE),
         coronavirus = na.approx(coronavirus, na.rm = FALSE),
         taste_loss  = na.approx(taste_loss, na.rm = FALSE),
         smell_loss  = na.approx(smell_loss, na.rm = FALSE))

@@

```{r suicide}
# suicide_measures_both <- read_delim(here('data','suicide',
#                                     'Multiple Cause of Death, 1999-2020.txt')) %>% 
#   janitor::clean_names() %>% 
#   filter(is.na(notes)) %>% 
#   select(fips = county_code, county, year = year_code, deaths_both = deaths, population)
# 
# suicide_measures_mcd <- read_delim(here('data','suicide',
#                                     'Multiple Cause of Death, 1999-2020 only multiple cause.txt')) %>% 
#   janitor::clean_names() %>% 
#   filter(is.na(notes)) %>% 
#   select(fips = county_code, county, year = year_code, deaths_mcd = deaths)
# 
# breaks <-  c(10,20,40,60,100,200,400,700,1300)
# suicide_measures <- suicide_measures_both %>% 
#   left_join(suicide_measures_mcd, by = c('fips', 'county', 'year')) %>% 
#   mutate(deaths_mcd = na_if(deaths_mcd,'Missing'),
#          deaths_both = na_if(deaths_both,'Missing'),
#          deaths_mcd_supp = ifelse(deaths_mcd == 'Suppressed', '(0:9]', NA), 
#          deaths_both_supp = ifelse(deaths_both == 'Suppressed', '(0:9]', NA), 
#          deaths_mcd = na_if(deaths_mcd,'Suppressed'),
#          deaths_both = na_if(deaths_both,'Suppressed'),
#          deaths_mcd_cut = as.character(cut(as.numeric(deaths_mcd), breaks = breaks)),
#          deaths_both_cut = as.character(cut(as.numeric(deaths_both), breaks = breaks)),
#          deaths_mcd_cut = ifelse(!is.na(deaths_mcd_supp),
#                                  deaths_mcd_supp,
#                                  deaths_mcd_cut), 
#          deaths_both_cut = ifelse(!is.na(deaths_both_supp),
#                                   deaths_both_supp,
#                                   deaths_both_cut)) %>% 
#         select(fips, county, year, deaths_both_cut) %>% 
#   mutate(deaths = factor(deaths_both_cut, ordered = T),
#          deaths = fct_relevel(deaths, c("(0:9]",
#                                         "(10,20]",
#                                         "(20,40]", 
#                                         "(40,60]",
#                                         "(60,100]",    
#                                         "(100,200]",
#                                         "(200,400]",
#                                         "(400,700]",   
#                                         "(700,1.3e+03]")))
# suicide_measures %>% 
#   count(year, deaths) %>%
#   ggplot(aes(x = deaths, y = n)) +
#   geom_bar(stat = 'identity') +
#   facet_wrap(~year)

suicide_measures <- read_delim(here('data','suicide',
                                    'Multiple Cause of Death, 1999-2020 only multiple cause.txt'),
                               show_col_types = FALSE) %>% 
  janitor::clean_names() %>% 
  filter(is.na(notes)) %>% 
  select(fips = county_code, county, year = year_code, deaths, population) %>% 
  mutate(across(c(deaths, population), ~na_if(.x,'Missing')),
         deaths = ifelse(deaths == 'Suppressed', '9', deaths), 
         population = na_if(population,'Suppressed'),
         across(c(deaths, population), as.numeric),
         death_rate = (deaths/ population) * 1000)

@@

suicide <- suicide %>% 
  group_by(fips) %>%
  mutate(deaths           = na.approx(deaths, na.rm = FALSE),
         population       = na.approx(population, na.rm = FALSE),
         suicide          = na.approx(suicide, na.rm = FALSE),
         depression       = na.approx(depression, na.rm = FALSE),
         suicide_hotline  = na.approx(suicide_hotline, na.rm = FALSE))

@@

suicide %>% 
  filter(fips %in% c("08003", '55057', '51059', '39085', 
           # '36119',
           '34021', '12071', '04017')) %>% 
  mutate(death_rate = death_rate * 100) %>% 
  pivot_longer(cols = c('suicide', 'depression', 'suicide_hotline', 'death_rate')) %>% 
  ggplot(aes(x = year, y = value, group = name, color = name)) +
  geom_line() + 
  geom_point() +
  facet_wrap(~ glue('{county}'), scale = 'free_y') + 
  theme_minimal()

### 2022-02-19 paper1.Rmd
write initial data and methods

---
title: "Article 1:	Construct Validity and Correspondence of Google Trends"
author: "Kelsey Gonzalez"
date: "`r format(Sys.time(), '%d %B, %Y')`"
editor_options:
  chunk_output_type: console
output:
  bookdown::pdf_document2:
    latex_engine: xelatex
    number_sections: true
  bookdown::word_document2: 
    toc: true
    reference_docx: StylesTemplate.docx
header-includes:
   - \usepackage{siunitx}
   - \usepackage{setspace}
   - \usepackage{dcolumn}
   - \usepackage{caption}
   - \usepackage{longtable}
   - \usepackage{booktabs}
   - \doublespacing
   - \usepackage{placeins}
csl: american-sociological-association.csl
bibliography: "thesis.bib"
indent: true
always_allow_html: true
---

@@

if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, glue, extrafont, here, jtools, psych,
               modelsummary, showtext, thematic, kableExtra, scales)

@@

conf_int_prop <- function(var){
  p <- mean(!!var, na.rm = T)
  q <- 1 - p
  n <- nrow(rain)
  min <- p - 1.96*sqrt((p*q)/n)
  max <- p + 1.96*sqrt((p*q)/n) 
  paste0(label_percent()(min), ", ", label_percent()(max))
}
print_perc <- function(var){
  label_percent()(mean(var, na.rm = T))
}

@@

# Intro 
New big data sources have led to vast possibilities for social science research
because they are bigger, cheaper, and already available
[@kingEnsuringDataRichFuture2011; @lazerComputationalSocialScience2009;
@salganikBitBitSocial2017]. Before overenthusiastically embracing these sources
into our workflows, social scientists must clearly establish parameters under
which these data sources could be operationalized
[@bailCulturalEnvironmentMeasuring2014; @lazerParableGoogleFlu2014]. As prior
research outlined, the “quantity of data does not mean that one can ignore
foundational issues of measurement and construct validity and reliability and
dependencies among data” [@lazerParableGoogleFlu2014, p. 1203]. Building on this
prior research outlining these various issues with big data
[@boydCriticalQuestionsBig2012; @lazerIssuesConstructValidity2015], this paper
tests the construct validity of Google Search Trends as an indicator of three
different cases, namely cultural attitudes, disease prevalence and voting
behavior. These three cases will be tested using cultural indicators from the
NORC General Social Survey, United States county-level suicide rates from
National Center for Health Statistics, US cases rates of Covid-19 from The New
York Times, historical US Presidential Election results, and the American
National Election Survey. Data will be analyzed against corresponding
operationalized Google Trends longitudinal data using Pearson’s r for pairwise
correlations, testing for the strength of relationships between the Google Trend
indicator and the respective comparison indicator. This paper will contribute to
the creation of methodological norms and standards of how to use Google Trends
as a big data source for societal research and serve as a critical inquiry into
the adoption of big data without a critical eye for the ecological validity of
the sources.

@@

I will use three categorizations of ways I propose Google Trends could be
operationalized for social scientific usage. First, I’ll test Google Trends as
an operationalization of cultural attitudes with the General Social Survey.
After @bailCulturalEnvironmentMeasuring2014’s call for cultural
sociologists to utilize the ever-expanding world of big data, Google trends as a
data source began appearing in sociological and social science research.  From
research on mass shootings and firearms [@brownsteinInternetSearchPatterns2020;
@semenzaInformationseekingWakeTragedy2020], protest and anti-Muslim sentiment
[@bailUsingInternetSearch2018; @barrieSearchingRacismGeorge2020;
@grossThereFergusonEffect2017], to analyzing country-level changes in social
perception [@reyes_etal18], Google search trends are a new and innovative
indicator of cultural interest. Extending into social networks and culture,
@bailPrestigeProximityPrejudice2019 even used Google
trends to measure how culture spreads around the globe.

Google Search Trends have also been used continuously in estimations of disease
prevalence and population health in journals like the Journal of Medical
Internet Research. While much of this research has focused on the Covid-19
pandemic [@jimenez_etal20; @jimenezCOVID19SymptomGoogle2020;
@limEstimatingInformationSeekingBehaviour2020;
@mavraganiCOVID19PredictabilityUnited2020; @nguyenGoogleTrendsAnalysis2020;
@todorovaInternetBasedData2021] , other research has investigated Google Trends
as an indicator of wellbeing [@brodeurCOVID19LockdownsWellbeing2021;
@carpiTwitterSubjectiveWellBeing2020; @duCOVID19IncreasesOnline2020],
suicidality [@burnettTimeTrendsPublic2020], vaccination uptake
[@dalumhansenEnsembleLearnedVaccination2016], obesity
[@sarigulNowcastingObesityUsing2014], and even insomnia
[@zittingGoogleTrendsReveal2020], to cover a few examples. For a partial
review of other utilizations, see @nutiUseGoogleTrends2014. According to
@jaidkaInformationseekingVsSharing2021, the majority of studies profess a
correlation of > .70, “demonstrating the vast potential of Google Search as a
proxy for monitoring population health” (p. 3) based on assumptions that
individuals search because of self-diagnosis and to identify possible courses
of treatment [@dechoudhurySeekingSharingHealth2014].

<!-- Following reports say that Google searches can predict covid  -->
<!-- 14. Brunori, P. & Resce, G. Searching for the peak Google Trends and the Covid-19 outbreak in Italy. SERIES working papers N. 04/2020 -->
<!-- Lampos, V. et al. Tracking COVID-19 using online search. Preprint at https://arxiv.org/abs/2003.08086 (2020). -->
<!-- 16. Walker, A., Hopkins, C. & Surda, P. The use of google trends to investigate the loss of smell related searches during COVID‐19 -->
<!-- outbreak. Int. Forum Allergy Rh. 10, 839–847 (2020). -->
<!-- 17. Stephens-Davidowitz, S. Google Searches Can Help Us Find Emerging Covid-19 Outbreaks. The New York Times https://www. -->
<!-- nytimes.com/2020/04/05/opinion/coronavirus-google-searches.html (2020). -->
<!-- 18. Frank, T. Goldman says fewer ‘loss of smell’ Google queries suggest better COVID outlook. Consumer News and Business Channel. CNBC https://www.cnbc.com/2020/04/15/goldman-says-fewer-loss-of-smell-google-searches-is-a-positive-sign-for-pande -->
<!-- mic.html (2020) -->

Various sources have also used Google Trends as a way to forecast political
elections and political attitudes [@wolfTrendingRightDirection2018]. For
instance, @swearingenGoogleInsightsSenate2014 investigate how U.S. Senate
Elections relate to attention measured by search traffic.
@prado-romanGoogleTrendsPredictor2020 compare how Google Search trends are able
to predict presidential election results in both the United States and Canada.
Finally, the OECD Development Centre is investigating how Google data can help
elucidate governments’ approval in Latin America [@montoyaUsingGoogleData2020].

@@

https://journals.sagepub.com/doi/10.1177/0894439316631043
			§ What constructs might google trends capture and not capture well? 
Capture attention but not attitudes

@@

# Research Methodology
To investigate the construct and criterion validity of the use of Google Trends
in these three areas, I will gather geo-located data from a sources related to
each case and use Pearson’s r for pairwise correlations, testing for the
strength of relationship between the Google Trend indicator and the respective
comparison indicator. Pairwise correlations allow for testing of nested data,
allowing for multiple search terms to be compared across locations
simultaneously as well as comparing change over time. I will test both raw
questions with raw trends as well as using dimensionality reduction through
factor analysis and principal components analysis (PCA). In addition, I’d like
to explore how this data would cluster locations together using tools like
blockmodeling and other clustering techniques.

To investigate the correlation between Google search trends, I will use
questions from the NORC General Social Survey. I will need to obtain an $800
grant to purchase geolocated data from NORC for this project. I plan to submit a
small grant application to SBS to receive these funds. I will use various NORC
GSS cultural value questions and aggregate the relevant zip code up to Nielsen’s
metropolitan market area (MMA), the level Google trends are aggregated to, using
population weighting.

For disease prevalence, I will use United States county-level suicide rates from
National Center for Health Statistics, US cases rates of Covid-19 from The New
York Times, and obesity rates to compare search trends to observable data.

Finally, I’ll test Google Trends as an operationalization of political attitudes
by looking at actual voting outcomes in historical US Presidential Election
results and the American National Election Survey and Google Trends about
candidates.

```{r}
tibble::tribble(
                                ~Validated.source, ~Google.trend.equivalent,
                                            "GSS",                       NA,
  "Suicide,\\nCovid-19 rates\\nVaccine hesitancy",                       NA,
           "Presidential Voting outcomes,\\nANES",                       NA
  )

@@

## Measures
### attitudinal
• Vaccine Hesitancy
		○ https://data.cdc.gov/Vaccinations/Vaccine-Hesitancy-for-COVID-19-County-and-local-es/q9mh-h2tw/data

• Mask Usage
		○ New York Times Survey

### health
• Covid Rates
		○ https://github.com/nytimes/covid-19-data/tree/master/rolling-averages

• Suicide
		○ https://wonder.cdc.gov/mcd.html 
		○ Multiple Cause of Death Data 2019 (see notes for which variables count)

### political
• Presidential results
		○ https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/VOQCHQ
		○ MIT Election Data and Science Lab, 2018, 
		○ "County Presidential Election Returns 2000-2020",
		○ https://doi.org/10.7910/DVN/VOQCHQ, 
		○ Harvard Dataverse, V9

### Independent Variables
## Analysis
# Results
## Cultural
## Medical
## Political
# Discussion
While I expect these tests to show high correlation between observed indicators
and Google search trends, there will be three important questions that surface.
First, just because something is correlated, does that mean it can replace the
collection of other types of data? Second, how correlated does a trend need to
be for social scientists to justifiably rely on it to indicate some outcome? And
finally, how can we construct analyses like this to be robust to changes in the
terms used across time and location?

@@

# References {-}

<div id="refs"></div>

>>>>

---
title: "Article 1:	Construct Validity and Correspondence of Google Trends"
author: "Kelsey Gonzalez"
date: "`r format(Sys.time(), '%d %B, %Y')`"
editor_options:
  chunk_output_type: console
  markdown: 
    wrap: 72
output:
  bookdown::pdf_document2:
    latex_engine: xelatex
    number_sections: true
  bookdown::word_document2: 
    toc: true
    reference_docx: StylesTemplate.docx
header-includes:
   - \usepackage{siunitx}
   - \usepackage{setspace}
   - \usepackage{dcolumn}
   - \usepackage{caption}
   - \usepackage{longtable}
   - \usepackage{booktabs}
   - \doublespacing
   - \usepackage{placeins}
csl: american-sociological-association.csl
bibliography: "thesis.bib"
indent: true
always_allow_html: true
---

@@

if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, glue, extrafont, here, jtools, psych,
               modelsummary, showtext, thematic, kableExtra, scales,
               rmcorr)

@@

```
```{r data}
vacc_hes <- read_rds(here('data', 'cleaned_data', 'vacc_hes.rds'))
mask <- read_rds(here('data', 'cleaned_data', 'mask.rds'))
covid <- read_rds(here('data', 'cleaned_data', 'covid.rds'))
suicide <- read_rds(here('data', 'cleaned_data', 'suicide.rds'))
pres_2016 <- read_rds(here('data', 'cleaned_data', 'pres_2016.rds'))
pres_2020 <- read_rds(here('data', 'cleaned_data', 'pres_2020.rds'))
```

@@

# Intro

New big data sources have led to vast possibilities for social science
research because they are bigger, cheaper, and already available
[@kingEnsuringDataRichFuture2011; @lazerComputationalSocialScience2009;
@salganikBitBitSocial2017]. Before overenthusiastically embracing these
sources into our workflows, social scientists must clearly establish
parameters under which these data sources could be operationalized
[@bailCulturalEnvironmentMeasuring2014; @lazerParableGoogleFlu2014]. As
prior research outlined, the "quantity of data does not mean that one
can ignore foundational issues of measurement and construct validity and
reliability and dependencies among data" \[@lazerParableGoogleFlu2014,
p. 1203\]. Building on this prior research outlining these various
issues with big data [@boydCriticalQuestionsBig2012;
@lazerIssuesConstructValidity2015], this paper tests the construct
validity of Google Search Trends as an indicator of three different
cases, namely cultural attitudes, disease prevalence and voting
behavior. These three cases will be tested using cultural indicators
from the NORC General Social Survey, United States county-level suicide
rates from National Center for Health Statistics, US cases rates of
Covid-19 from The New York Times, historical US Presidential Election
results, and the American National Election Survey. Data will be
analyzed against corresponding operationalized Google Trends
longitudinal data using Pearson's r for pairwise correlations, testing
for the strength of relationships between the Google Trend indicator and
the respective comparison indicator. This paper will contribute to the
creation of methodological norms and standards of how to use Google
Trends as a big data source for societal research and serve as a
critical inquiry into the adoption of big data without a critical eye
for the ecological validity of the sources.

@@

I will use three categorizations of ways I propose Google Trends could
be operationalized for social scientific usage. First, I'll test Google
Trends as an operationalization of cultural attitudes with the General
Social Survey. After @bailCulturalEnvironmentMeasuring2014's call for
cultural sociologists to utilize the ever-expanding world of big data,
Google trends as a data source began appearing in sociological and
social science research. From research on mass shootings and firearms
[@brownsteinInternetSearchPatterns2020;
@semenzaInformationseekingWakeTragedy2020], protest and anti-Muslim
sentiment [@bailUsingInternetSearch2018;
@barrieSearchingRacismGeorge2020; @grossThereFergusonEffect2017], to
analyzing country-level changes in social perception [@reyes_etal18],
Google search trends are a new and innovative indicator of cultural
interest. Extending into social networks and culture,
@bailPrestigeProximityPrejudice2019 even used Google trends to measure
how culture spreads around the globe.

Google Search Trends have also been used continuously in estimations of
disease prevalence and population health in journals like the Journal of
Medical Internet Research. While much of this research has focused on
the Covid-19 pandemic [@jimenez_etal20;
@jimenezCOVID19SymptomGoogle2020;
@limEstimatingInformationSeekingBehaviour2020;
@mavraganiCOVID19PredictabilityUnited2020;
@nguyenGoogleTrendsAnalysis2020; @todorovaInternetBasedData2021] , other
research has investigated Google Trends as an indicator of wellbeing
[@brodeurCOVID19LockdownsWellbeing2021;
@carpiTwitterSubjectiveWellBeing2020; @duCOVID19IncreasesOnline2020],
suicidality [@burnettTimeTrendsPublic2020], vaccination uptake
[@dalumhansenEnsembleLearnedVaccination2016], obesity
[@sarigulNowcastingObesityUsing2014], and even insomnia
[@zittingGoogleTrendsReveal2020], to cover a few examples. For a partial
review of other utilizations, see @nutiUseGoogleTrends2014. According to
@jaidkaInformationseekingVsSharing2021, the majority of studies profess
a correlation of \> .70, "demonstrating the vast potential of Google
Search as a proxy for monitoring population health" (p. 3) based on
assumptions that individuals search because of self-diagnosis and to
identify possible courses of treatment
[@dechoudhurySeekingSharingHealth2014].

<!-- Following reports say that Google searches can predict covid  -->

<!-- 14. Brunori, P. & Resce, G. Searching for the peak Google Trends and the Covid-19 outbreak in Italy. SERIES working papers N. 04/2020 -->

<!-- Lampos, V. et al. Tracking COVID-19 using online search. Preprint at https://arxiv.org/abs/2003.08086 (2020). -->

<!-- 16. Walker, A., Hopkins, C. & Surda, P. The use of google trends to investigate the loss of smell related searches during COVID‐19 -->

<!-- outbreak. Int. Forum Allergy Rh. 10, 839–847 (2020). -->

<!-- 17. Stephens-Davidowitz, S. Google Searches Can Help Us Find Emerging Covid-19 Outbreaks. The New York Times https://www. -->

<!-- nytimes.com/2020/04/05/opinion/coronavirus-google-searches.html (2020). -->

<!-- 18. Frank, T. Goldman says fewer ‘loss of smell’ Google queries suggest better COVID outlook. Consumer News and Business Channel. CNBC https://www.cnbc.com/2020/04/15/goldman-says-fewer-loss-of-smell-google-searches-is-a-positive-sign-for-pande -->

<!-- mic.html (2020) -->

Various sources have also used Google Trends as a way to forecast
political elections and political attitudes
[@wolfTrendingRightDirection2018]. For instance,
@swearingenGoogleInsightsSenate2014 investigate how U.S. Senate
Elections relate to attention measured by search traffic.
@prado-romanGoogleTrendsPredictor2020 compare how Google Search trends
are able to predict presidential election results in both the United
States and Canada. Finally, the OECD Development Centre is investigating
how Google data can help elucidate governments' approval in Latin
America [@montoyaUsingGoogleData2020].

@@

<https://journals.sagepub.com/doi/10.1177/0894439316631043> § What
constructs might google trends capture and not capture well? Capture
attention but not attitudes

@@

# Research Methodology
To investigate the construct and criterion validity of the use of Google Trends
in these three areas, I gathered geo-located social science data across multiple
sources to address the three areas of inquiry for this paper.  Table
\@ref(tab:data-sources-table) outlines which data sources are used for this
project and which trends are matched to each source.

```{r data-sources-table}
#| results='asis',
#| tab.cap = 'airquality dataset caption here'

data_sources <- tibble::tribble(
~Validated.Data.Source, ~Type, ~Dates, ~Google.Trends.Used,
"General Social Survey", "Cross-Sectional", "2010 - 2020", NA,

"Vaccine Hesitancy for COVID-19", "Cross-Sectional", "March 3 – 15, 2021", "covid conspiracy', 'COVID-19 vaccine', 'Coronavirus', 'Covid-19'",

"Mask-Wearing Survey Data", "Cross-Sectional", "July 2 - 14, 2020", "'Face Mask', 'Mask', 'Cloth Face Mask'",

 "Covid Rates", "Longitudinal", "Every Monday, 2020 - 2021", "'Covid-19', 'Coronavirus', 'Taste Loss', 'Smell Loss'",

"County Suicide Rates", "Longitudinal", "Yearly 2010-2020", "'Suicide', 'Depression',' Suicide Hotline'",

"American National Election Survey", "Cross-Sectional", "2020", NA,

"Presidential Election Results", "Cross Sectional", "2016 & 2020", "'Hilary Clinton', 'Donald Trump', 'Joe Biden'"
 )

data_sources %>% 
  kbl() %>%
  # group rows
  pack_rows("Behaviors and Attitudes", 1,3, background = '#bad7db', bold = T) %>% 
  pack_rows("Health", 4,5, background = '#bad7db', bold = T) %>% 
  pack_rows('Political', 6,7, background = '#bad7db', bold = T) %>% 
  # header style
  row_spec(0, background = my_palette[8], color = 'white', bold = T) %>% 
  #overall styling
  kable_styling(latex_options = c( "hold_position"), full_width = T)

@@

## Measures
### Google Trends
This paper focuses on a validation of Google search trends [@googletrends].
Google trends portray the search frequency for specific search terms across
designated media markets areas (DMAs), a nonoverlapping aggregation of U.S.
counties to 210 media markets based on similar population clusters [@dma_key].
Raw data is on a scale from 0 to 100, with 100 being the maximum search
popularity out of all DMAs. Google Search Trend are now only available
cross-sectionally (a single time period across a geography) or as time-series (a
single geo-location across time). To remedy this and build a longitudinal
dataset of each search topic for the longitudinal datasets, I follow the method
proposed in @park_etal (p. 5). This method involves building a dataset of
unscaled cross-sectional values, selecting a DMA to use to establish the
rescaling ratio (I use 'Los Angeles CA'), and then finding the time-series
values for the one DMA. To find the rescaling ratio for each week in the
time-series, you divide the time-series value for each week by the
cross-sectional value for each week, resulting in a rescaling vector to be used
for all weeks in the dataset across geographies. To rescale each longitudinal
value, multiply the respective week's rescaling ratio by the cross-sectional
value. Rescaled longitudinal data was compared against time-series data for
multiple test counties and was equivalent. For a more in-depth explanation of
this procedure, see @park_etal (p. 5). Missing datapoints in longitudinal
datasets were filled in with interpolated values using `zoo::na.approx()`
[@zoo].

### attitudinal

<!-- [@gss_data] (let's see if I get access first) -->

One measure of attitudinal indicators is the Vaccine Hesitancy for COVID-19
[@vaches_data]. The CDC uses the U.S. Census Bureau’s Household Pulse Survey
(HPS) and the 2019 American Community Survey (ACS) 1-year Public Use Microdata
Sample (PUMS) to measure U.S. residents’ intentions to receive the COVID-19
vaccine if available during May 26, 2021 – June 7, 2021. This dataset consists
of `r nrow(vacc_hes)` observations, one for each U.S. County.

Another attitudinal indicator I use is the Mask-Wearing Survey Data conducted by
Dynata for the New York Times from July 2 through 14, 2020 [@mask_data]. 250,000
survey respondents were asked, "How often do you wear a mask in public when you
expect to be within six feet of another person?". The NYT weighted each response
to create a county level measure of what percent of the county never, rarely,
sometimes, frequetly, and always wore a mask when in public. This dataset
consists of `r nrow(mask)` observations, one for each U.S. County.

### health

I first use U.S. Covid-19 rates to validate health and disease related topics. I
retrieve U.S. county-level Covid-19 rates from by @covid_data, who compile this
data based on reports from state and local health agencies. It is widely
acknowledged that there are biases in this data due to inconsistencies and
availability in testing as well as different community propensity to test
[@gu22; @cdc20a] However, it is the best measure we have of actual case rates.
Case Rate is measured as number of cases per 100,000 population. Observations
vary from a minimum of 0 to a maximum of `r max(covid$covid_rate))` for each 
Monday from `r glue("{format( min(covid$date), '%B %d, %Y')} through {format(max(covid$date), '%B %d, %Y')}")`.
There are `r nrow(covid)` cases across `r nrow(count(covid, fips))` 
counties and `r nrow(count(covid, date))` dates. Missing data were interpolated 
using `zoo::na.approx()` [@zoo].

I also use county-level suicide rates from the US Centers for Disease Control
and Prevention [-@suic_data]. Data is grouped by year from 2010-2020. Raw death
rates are scaled by population size for each year. There are `r nrow(suicide)` 
total cases, resulting from `r nrow(count(suicide, year))` observations of 
`r nrow(count(suicide, fips))` counties. Missing data were interpolated 
using `zoo::na.approx()` [@zoo].

### political
Finally, I test Google Trends as an operationalization of political
attitudes by first looking at actual voting outcomes in historical US
Presidential Election results. This data comes from @pres_data, who scraped the
results from Townhall.com, Fox News, Politico, and the New York Times. Data on 
presidential outcomes were avaiable for `r nrow(pres_2016)` counties in 2016 and
`r nrow(pres_2020)` counties in 2020.

In addition, I use data on political opinions from the the American National
Election Survey 2020 Time-Series Study [@anes_data]. This data is paired with
restricted data provided by ICPSR, the Inter-university Consortium for Political
and Social Research, that contains the geo-ids of the `r 5,441` survey respondents.
The study interviewed respondents in a pre-election survey that was conducted
between August 18, 2020 and the day of the US Presidential election day,
November 3, 2020.

## Analysis
For cross-sectional numeric data, I use the Pearson correlation formula (formula
\@ref(eq:pearsoncorr)) to calculate the strength of the relationship between
each Google Trend and the respective datasource.

\begin{equation}
 r =
  \frac{ \sum_{i=1}^{n}(x_i-\bar{x})(y_i-\bar{y}) }{
        \sqrt{\sum_{i=1}^{n}(x_i-\bar{x})^2}\sqrt{\sum_{i=1}^{n}(y_i-\bar{y})^2}} (\#eq:pearsoncorr)
\end{equation}

To address longitudinal correlations, I employ Repeated Measures Correlation
[@bland1995] using the `rmcorr` package in R [@bakdash2017]. Repeated Measures
Correlation is useful for determining the within-individual association for
paired measures across time and individuals.

I test both the raw trends, as well a combination of trends created
through factor analysis as a method of dimensionality reduction.

As an additional test of correlation, I employ multiple linear
regression to identify the strength of relationships across locales: I
include city-data like county population size, internet usage rates, and
Z to attempt to disentangle why some individual cases may be better predicted by
Google Trends while others may not.

visualizations made with [@wickham11; @wickham_etal]

# Results

## Cultural

## Medical

```{r}
covid_corr <- rmcorr::rmcorr(fips, measure1 = covid_rate, measure2 = covid_19, dataset = covid)

covid_reg <- lm(covid_rate ~ covid_19 + smell_loss + taste_loss, data = covid)

```{r}
suicide_corr <- rmcorr::rmcorr(fips, measure1 = death_rate, measure2 = suicide, dataset = suicide)

suicide_reg <- lm(death_rate ~ suicide + depression + suicide_hotline, data = suicide)
summary(suicide_reg)

library(nlme)

suicide <- suicide %>% 
  drop_na()
suicide_lme <- lme(fixed = death_rate ~ suicide + depression + suicide_hotline,
             random = ~ 1 | fips,
             data = suicide)
summary(suicide_lme)
```

## Political

# Discussion

While I expect these tests to show high correlation between observed
indicators and Google search trends, there will be three important
questions that surface. First, just because something is correlated,
does that mean it can replace the collection of other types of data?
Second, how correlated does a trend need to be for social scientists to
justifiably rely on it to indicate some outcome? And finally, how can we
construct analyses like this to be robust to changes in the terms used
across time and location?

@@

# References {.unnumbered}

::: {#refs}
:::
